\documentclass[aspectratio=32]{beamer}


% Load packages
\usepackage{datetime}
\usepackage{ifthen}
\usepackage{ragged2e}
\usepackage{fancyvrb}   % Alternative verbatim environment
\usepackage{graphicx}      % For handling graphics
\usepackage{amssymb}       % For mathematical symbols
\usepackage{amsmath}
\usepackage{xparse}
\usepackage{pgfplotstable} % For handling CSV data
\usepackage{fancyhdr}
\usepackage{parskip} % Removes default interparagraph spacing
\usepackage{booktabs}      % For publication quality tables
\usepackage{siunitx}       % For handling units and numbers
\usepackage{listings}      % For source code listings
\usepackage{caption}       % For captions customization
\usepackage{etoolbox}      % For programming tools
\usepackage{hyperref}      % For cross-referencing and hyperlinks
\usepackage{float}         % For floating environments
\usepackage{multicol}      % For multi-column layout
\usepackage{longtable}     % For long tables
\usepackage{adjustbox}     % For adjusting boxes
\usepackage{xcolor}        % For color customization
\usepackage[backend=biber, style=numeric]{biblatex}

% code listing
\lstset{
    basicstyle=\ttfamily, % Use monospace font at a smaller size
    breaklines=true, % Enable line breaking
    numbers=left, % Adjust or remove according to preference
    tabsize=4, % Reduce tab size for more compact code
    showstringspaces=true, % Don't show spaces in strings as special characters
    frame=single, % Optional: Adds a frame around the code block
    framesep=2pt, % Reduce frame separation
    aboveskip=3pt, % Reduce space above the code block
    belowskip=3pt, % Reduce space below the code block
    xleftmargin=5pt, % Reduce left margin
    xrightmargin=5pt % Reduce right margin
}



% Metadata
\title{Mechanistic Interpretability on Irreducible Integer Identifiers}
\date{\mydate\today}
\author{Noah Syrkis}
\newcommand{\affiliation}{University of
Copenhagen} % Capture affiliation from the header
\newcommand{\footerurl}{false} % Capture the URL from the header

% BibTeX setup
\addbibresource{/Users/syrkis/code/press/library.bib}

% Listings and date setup
\newdateformat{mydate}{\shortmonthname[\the\month]. \THEDAY, \THEYEAR}
% tight list
\providecommand{\tightlist}{\setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}

% Font setup
\usefonttheme{serif}
\renewcommand{\baselinestretch}{1.5}

% Text color setup
\setbeamercolor{normal text}{fg=black,bg=white}
\setbeamercolor{frametitle}{fg=black,bg=white}
\setbeamercolor{title}{fg=black,bg=white}
\setbeamercolor{bibliography entry author}{fg=black}
\setbeamercolor{bibliography entry title}{fg=black}
\setbeamercolor{bibliography entry location}{fg=black}
\setbeamercolor{bibliography entry note}{fg=black}
\setbeamercolor{bibliography entry journal}{fg=black}
\setbeamercolor{bibliography item}{fg=black}
\setbeamercolor{tableofcontents}{fg=black}
\setbeamercolor{section in toc}{fg=black}
\setbeamercolor{subsection in toc}{fg=black}
\setbeamercolor{itemize item}{fg=black}
\setbeamercolor{item}{fg=black}
\setbeamercolor{caption name}{fg=black}
\setbeamercolor{caption}{fg=black}

% Miscellaneous settings
\setbeamertemplate{frametitle continuation}[from second]
% section in toc number should be without dot after number
\setbeamertemplate{section in toc}{\inserttocsectionnumber{} \inserttocsection}
\setbeamertemplate{navigation symbols}{}
\setbeamertemplate{caption}[numbered]
\setbeamertemplate{headline}{}

% Footer setup
\setbeamertemplate{footline}
{
  \leavevmode
  \hbox{
    \ifnum\thepage>1
      \hspace{0.4cm}\insertframenumber{} of \inserttotalframenumber
      \begin{beamercolorbox}[wd=\paperwidth, ht=2.5ex,dp=1ex]{page number}
        % Check if URL is defined and not set to "false"
        \ifthenelse{\equal{\footerurl}{} \OR \equal{\footerurl}{false}}{}{%
          \hfill\footerurl\hspace{1.7cm}
        }
      \end{beamercolorbox}
    \fi
  }
    \vspace*{0.4cm}
}

% Modify the frametitle template
\setbeamertemplate{frametitle}
{
  \vskip0.4cm
  \leavevmode
  \hbox{%
    \begin{beamercolorbox}[wd=\paperwidth,ht=0ex,dp=0ex]{frametitle}
      \hskip0.7cm\usebeamerfont{frametitle}%
        {\insertsectionnumber~\insertframetitle} % Otherwise, display the frame title with section number
    \end{beamercolorbox}
  }
}

% \AtBeginSection[]{
%   \begin{frame}[allowframebreaks]
%     \vfill
%     \tableofcontents[hideallsubsections]
%     \vfill
%   \end{frame}
% }


% Custom title page
\defbeamertemplate*{title page}{customized}
{
  \begin{minipage}[c][\textheight][c]{0.5\textwidth}
    \begin{center}
      \usebeamerfont{title}\inserttitle\par
      \vspace{1cm}
      \small\usebeamerfont{author}\insertauthor\par
      \ifthenelse{\equal{\affiliation}{}}{}{% Check if affiliation is defined
        \small\affiliation\par % Display affiliation
      }
      \small\usebeamerfont{date}\insertdate\par
    \end{center}
  \end{minipage}%
  \hfill
  \begin{minipage}[c][0.7\textheight][c]{0.4\textwidth}
    \tableofcontents[hideallsubsections]
  \end{minipage}
}

\newlength{\maxwidth}
\setlength{\maxwidth}{.70\textwidth} % Example: 80% of the text width
\newlength{\maxheight}
\setlength{\maxheight}{.70\textheight} % Example: 70% of the text height
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}

\begin{document}

  % Front page
  \ifthenelse{\equal{\inserttitle}{}}{}{% Check if title is empty
  \begin{frame}[allowframebreaks]
    \titlepage
  \end{frame}
  }

  \begin{frame}[allowframebreaks]
  ``This disgusting pile of matrices with some non-linearities in
  between is an impressively poorly written beautiful and concise
  algorithm''\footnote<.->{Neel Nanda (not verbatim)}
  \end{frame}

  \section{\textbar{} Mech. interp. (MI)}\label{mech.-interp.-mi}

  \begin{frame}[allowframebreaks]{\textbar{} Mech. interp. (MI)}
  \begin{itemize}
  \tightlist
  \item
    To look into black (or opaque) machine learning (ML) boxes.
  \item
    Reverse-engineering deep learned circuits.
  \item
    \textcite{nanda2023} shows MI on modular addition transformer.
  \item
    We show MI on prime number classification transformer.
  \end{itemize}
  \end{frame}

  \subsection{\textbar{} On complexity}\label{on-complexity}

  \begin{frame}{\textbar{} On complexity}
  \begin{itemize}[<+->]
  \tightlist
  \item
    ML complexity (probably) grows faster than MI \autocite{sarkar2022}.
  \item
    Lessons from MI might still inform ML development and risk.
  \item
    Work on automatic MI \autocite{conmy2023}: ML might have a rôle in
    MI.
  \item
    Current ML is sub-symbolic (lacks the rigor of formal systems).
  \item
    Metric frenzy has made ML more engineering than science.
  \end{itemize}
  \end{frame}

  \subsection{\textbar{} Grokking}\label{grokking}

  \begin{frame}[allowframebreaks]{\textbar{} Grokking}
  \begin{itemize}
  \tightlist
  \item
    When a model suddenly generalizes \autocite{power2022}.
  \item
    Grokking means the weights represent an algorithm \ldots{}
  \item
    \ldots{} rather than a dataset.
  \item
    Good for MI, as it means circuits are there to be discovered.
  \end{itemize}
  \end{frame}

  \section{\textbar{} Transformers}\label{transformers}

  \begin{frame}[allowframebreaks]{\textbar{} Transformers}
  \begin{itemize}
  \tightlist
  \item
    Famously introduced by \textcite{vaswani2017}.
  \item
    Batch normalization, residual streams, projections, etc.
  \item
    We use \textcite{he2023}'s simplified transformer block.
  \end{itemize}
  \end{frame}

  \begin{frame}[allowframebreaks]
  ``God made the natural numbers; all else is the work of
  man.''\footnote<.->{Leopold Kronecker (also not verbatim)}
  \end{frame}

  \section{\texorpdfstring{\textbar{} \(\mathbb{Z}\)}{\textbar{} \textbackslash mathbb\{Z\}}}\label{mathbbz}

  \begin{frame}[allowframebreaks]{\textbar{} \(\mathbb{Z}\)}
  \begin{columns}[c]
  \begin{column}{0.6\textwidth}
  \small

  \begin{itemize}
  \tightlist
  \item
    Complexity from trivial in seq. \ref{eq:nat} to impossible in seq.
    \ref{eq:bb} (busy beaver \autocite{aaronson2020}).
  \item
    OEIS \autocite{sloane2003} is a big database of \(\mathbb{Z}\)-seqs.
  \item
    Four \(\mathbb{Z}\) tasks: classify, compare, continue, and unmask
    \autocite{belcak2022}.
  \item
    We focus on primes (seq. \ref{eq:prime}), which \textcite{lee2024}
    shows is doable.
  \end{itemize}
  \end{column}

  \begin{column}{0.3\textwidth}
  \begin{equation}\phantomsection\label{eq:nat}{
  0,  1, 2, 3, \dots
  }\end{equation}

  \begin{equation}\phantomsection\label{eq:bb}{
  6, 21, 107, \dots
  }\end{equation}

  \begin{equation}\phantomsection\label{eq:prime}{
  2, 3, 5, 7, 9, \dots
  }\end{equation}
  \end{column}
  \end{columns}
  \end{frame}

  \subsection[\textbar{} Irreducible
  integers]{\texorpdfstring{\textbar{} Irreducible
  integers\footnote<.->{\tiny Is what you call prime numbers when you
    really want the acronym of your project title to be MIIII.}}{\textbar{} Irreducible integers}}\label{irreducible-integers}

  \begin{frame}[allowframebreaks]{\textbar{} Irreducible integers\footnote<.->{\tiny Is
    what you call prime numbers when you really want the acronym of your
    project title to be MIIII.}}
  \begin{itemize}
  \tightlist
  \item
    Given a sequence from \(\mathbb{Z}\), which numbers are prime?
  \item
    Tests to determine primality include:

    \begin{itemize}
    \tightlist
    \item
      Wilson's Theorem: \(n>1\) is prime if \((n-1)\equiv -1 \mod n\).
    \item
      Fermat's Little Theorem: \(a^{n-1} \mod n = 1\), for \(a < n\).
    \item
      Euler's Criterion, AKS Primality Test, Miller-Rabin Primality
      Test, and more.
    \end{itemize}
  \end{itemize}
  \end{frame}

  \subsection{\textbar{} Irreducible integers
  (cont.)}\label{irreducible-integers-cont.}

  \begin{frame}[allowframebreaks]{\textbar{} Irreducible integers (cont.)}
  \begin{longtable}[]{@{}llllrrrr@{}}
  \caption{Four digit dataset with numbers and labels
  (\([ \textbf{X} | \textbf{Y} ]\)).}\tabularnewline
  \toprule\noalign{}
  \(x_0\) & \(x_1\) & \(x_2\) & \(x_3\) & \(y_0\) & \(y_1\) & \(y_2\) &
  \(y_3\) \\
  \midrule\noalign{}
  \endfirsthead
  \toprule\noalign{}
  \(x_0\) & \(x_1\) & \(x_2\) & \(x_3\) & \(y_0\) & \(y_1\) & \(y_2\) &
  \(y_3\) \\
  \midrule\noalign{}
  \endhead
  1001 & 1003 & 1007 & 1009 & 0 & 0 & 0 & 1 \\
  1011 & 1013 & 1017 & 1019 & 0 & 1 & 0 & 1 \\
  \(\vdots\) & & & & & & & \(\vdots\) \\
  9981 & 9983 & 9987 & 9989 & 0 & 0 & 0 & 0 \\
  9991 & 9993 & 9997 & 9999 & 0 & 0 & 0 & 0 \\
  \bottomrule\noalign{}
  \end{longtable}
  \end{frame}

  \section{\texorpdfstring{\textbar{} The MU puzzle
  \autocite{hofstadter2000}}{\textbar{} The MU puzzle {[}@hofstadter2000{]}}}\label{the-mu-puzzle-hofstadter2000}

  \begin{frame}[fragile,allowframebreaks]{\textbar{} The MU puzzle
  \autocite{hofstadter2000}}
  \begin{itemize}
  \tightlist
  \item
    Can you get \texttt{MI} from
    \texttt{MU} by:

    \begin{enumerate}
    \tightlist
    \item
      Adding a~\texttt{U}~to the end of any string
      ending in~\texttt{I}.
    \item
      Doubling the string after the \texttt{M}.
    \item
      Replacing any~\texttt{III}~with
      a~\texttt{U}.
    \item
      Removing any \texttt{UU}.
    \end{enumerate}
  \end{itemize}
  \end{frame}

  \section{\textbar{} MIIII}\label{miiii}

  \begin{frame}[allowframebreaks]{\textbar{} MIIII}
  \begin{itemize}
  \tightlist
  \item
    We something soemthign
  \item
    Then this
  \item
    Then this
  \item
    Then this
  \end{itemize}

  \framebreak

  \begin{itemize}
  \tightlist
  \item
    But then this.
  \end{itemize}
  \end{frame}

  \subsection{\textbar{} MIIII (cont.)}\label{miiii-cont.}

  \begin{frame}[fragile]{\textbar{} MIIII (cont.)}
\begin{lstlisting}[title={The algorithm}]
def fib(n):
    | -> return n :: []
    | Match P <- <> DIV (1 \in N)
\end{lstlisting}
  \end{frame}

  % bibliography

\begin{frame}[allowframebreaks,allowframebreaks]
  \Large{References}
  \small\linespread{1.2}\printbibliography
\end{frame}

\end{document}
